Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by [omitted for blind review] {\bf Note that all references to appendices will be replaced with URLs in the final paper; they are appendices only for blind review.}}

### 2022-10-12 paper/Paper.tex
Added WEST algorithm to section4, rewrote/removed parts of theoretical complexity, more efficient formatting in section7, lots of touchingup of appendix format - Chiara, Laura, Zili

%KYR: Take this out and replace it with a statement that we use the standard RE operators:...
%% \begin{definition}[\emph{Regular Expression \cite{sipser}}]
%% A \emph{regular expression} is a string that describes the form of a language. 
%% Let $\Sigma$ denote a finite alphabet. The empty set $\emptyset$, the empty string $\epsilon = ``"$, and any character in $\Sigma$ are constants defined to be a regular expression. 
%% Beyond that, we define regular expressions recursively. 
%% Suppose $R$ and $T$ are regular expressions. 
%% Then the following operations generate regular expressions:
%% \begin{itemize}
%%     \item \emph{(Concatenation)} Denoted $RT$, this operation describes the set of strings obtained by concatenating any string generated by $R$ and any string generated by $T$ in that order.
%%     \item \emph{(Alternation)} Denoted $R | T$, this operation describes the set union of the strings generated by $R$ and $T$.
%%     \item \emph{(Kleene Star)} Denoted $R^*$, this operation describes the set of all strings made by concatenating any finite number (including zero) of strings described by $R$ together.
%%     \item $R^i$ denotes regular expression consisting of $R$ repeated $i$ times for $i \geq 0$. $R^0 = \epsilon$.
%% \end{itemize}
%% \end{definition}
%
%KYR: Examples are super nice, but if we have them we should have them after the MLTL regex example
%\noindent We provide an example of each operation.
%% \begin{example}
%% Let $R$ describe \{`$a$', `$b$'\} and $T$ describe \{ `$c$', `$d$' \}.
%% \begin{itemize}
%%     \item $RT$ describes the set \{ `$ac$', `$ad$', `$bc$', `$bd$' \}.
%%     \item $R | T$ describes the set \{`$a$', `$b$', `$c$', `$d$' \}.
%%     \item $R^*$ describes the set \{ $\epsilon$, `$a$', `$b$', `$aa$', `$bb$', `$ab$', `$ba$', `$aaa$', `$aba$',... \}.
%%     \item $R^2$ describes the set \{ `$aa$', `$bb$', `$ab$', `$ba$' \}.
%% \end{itemize}
%% \end{example}

%% %\noindent We provide examples of different regex and possible combinations of the above operations.
%% \begin{example}
%% Let `$a$', `$b$', and `$c$' denote characters in a finite alphabet.
%% \begin{itemize}
%%     \item $ab^*$ describes the set \{ `$a$', `$ab$', `$abb$', `$abbb$', ... \}.
%%     \item $a^*|b$ describes the set \{ $\epsilon$, `$b$', `$a$', `$aa$', `$aaa$', ...\}.
%%     \item $a(b^*)c$ describes the set \{ `$ac$', `$abc$', `$abbc$', `$abbbc$', ... \}.
%%     \item $(a | b)^3$ describes the set \{ `$aaa$', `$aba$', `$aab$', `$abb$', `$bbb$', ... \}.
%% \end{itemize}
%% \end{example}

@@

 \end{itemize}

% \vspace{-0.1in}
% \subsection{Soundness and Completeness}
%  \vspace{-0.1in}
 We write $\mathscr{L}(\text{reg}(\phi))$ to denote the language of reg($\phi$). That is, the set of computations represented by the regular expression $\text{reg}(\phi)$. The proof of Theorem \ref{SoundComplete} appears in Appendix \ref{Appendix-SoundComplete}.

@@

%%  \noindent This completes the inductive step, and thus the proof. Since this proof addresses all possible MLTL formulas in negation normal form, it shows completeness along with soundness.
%%  \end{proof}
% \newpage
%\vspace{-0,1in}
%\subsection{Nested Until and Release Rewriting Theorem}
%\vspace{-0,1in}
 We prove the correctness of our formula rewriting algorithm using our regular expressions from Definition \ref{TRE}. As detailed in Section \ref{complex}, nested binary temporal operators create the worst-case time and space complexity for the WEST program. Although the WEST program does not implement this theorem by default, we would expect selected applications of rewriting to improve runtime. We use this theorem to simplify MLTL formulas, making them easier to interpret.
Theorem \ref{NestedUR}'s proof appears in Appendix \ref{Appendix-NestedUR}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \vspace{-0.1in}
 \section{WEST Code Specifications} \label{specs}
  \vspace{-0.1in}
The code for the WEST program can be found on its %\href{https://github.com/zwang271/2022-Iowa-State-REU-Temporal-Logic-.git} KYR: blind review
{GitHub} repository. The MLTL syntax for our program is given as the context free grammar below. In the alphabet, $T$ and $!$ refer to the propositional constants `true' and `false', respectively. $\&$, $v$, $=$, $>$ refer to propositional logic operators `and', `or, `equivalent', and `implies', respectively. Note that the last comma listed is not a typo and indeed part of the alphabet. \\
\textbf{Alphabet}: 
\begin{align*}
    \Sigma = \{&\text{`0' , `1' , `2' , `3' , `4' , `5' , `6' , `7' , `8' , `9' , `T' , `!' , `p' , `$\scriptstyle{\sim}$' , `v' , `\&' , `=' , `\mbox{$>$}' ,}\\
    &\text{ `F' , `G' , `U', `R' , `(' , `)' , `[' , `]' , `:' , `,'}\}
\end{align*}

\begin{align*}
    \text{Digit} &\to 0 \,|\, 1 \,|\,2 \,|\,3 \,|\,4 \,|\,5 \,|\,6 \,|\,7 \,|\, 8 \,|\, 9 \\
    \text{Num} &\to \text{Digit} \; \text{Num} \,|\, \text{Digit} \\
    \text{Interval} &\to [ \; \text{Num} \; : \; \text{Num} \; ] \\
    \text{Prop\_cons} &\to \text{T} \;|\; ! \\
    \text{Prop\_var} &\to \text{p} \; \text{Num} \\
    \text{Unary\_Prop\_conn} &\to \hspace{1mm}\scriptstyle{\sim} \\
    \text{Binary\_Prop\_conn} &\to \text{v} \;|\; \& \;|\; = \;|\; > \\
    \text{Assoc\_Prop\_conn} &\to \text{v} \;|\; \& \;|\; = \\
    \text{Array\_entry} &\to \text{Wff} \; , \; \text{Array\_entry} \;|\; \text{Wff} \\
    \text{Unary\_Temp\_conn} &\to \text{F} \;|\; \text{G} \\
    \text{Binary\_Temp\_conn} &\to \text{U} \;|\; \text{R} \\
    \text{Wff} &\to \text{Prop\_cons} \;|\; \text{Prop\_var} \\
               &\hspace{5mm} \;|\; \text{Unary\_Prop\_conn} \;\; \text{Wff} \\
               &\hspace{5mm} \;|\; \text{Wff} \;\; \text{Binary\_Prop\_conn} \;\; \text{Wff}\\
               &\hspace{5mm} \;|\; ( \; \text{Assoc\_Prop\_conn} \; [ \; \text{Array\_entry} \; ] \; ) \\
               &\hspace{5mm} \;|\; \text{Unary\_Temp\_conn} \;\; \text{Interval} \;\; \text{Wff} \\
               &\hspace{5mm} \;|\; ( \; \text{Wff} \;\; \text{Binar\_Temp\_conn} \;\; \text{Interval} \;\; \text{Wff} \; )
\end{align*}

In the grammar, associative propositional connectives (\textit{Assoc\_Prop\_conn}) are included for the ease of the user. Given well-formed formulas $\phi_1, \phi_2, \phi_3$, the following formula in prefix notation $(\lor[\phi_1, \phi_2, \phi_3])$ is parsed into $((\phi_1 \lor \phi_2) \lor \phi_3)$, which is in turn equivalent to $(\phi_1 \lor \phi_2 \lor \phi_3)$. \\
\begin{remark}
For two propositional variables, the associative equivalence operator functions identically to the binary propositional connective equivalence operator. For formulas with 3 or more propositional variables, however, the associative equivalence operator does not mean that each propositional variable in the list is equivalent. Instead, it means that the equivalence of the first two propositional variables in the list is equivalent to the next propositional variable in the list, and the truth value for this expression is equivalent to the next propositional variable, and so on. For example, $=[p1,p2,p3]$ is equivalent to $(p1=p2)=p3$ rather than $p1=p2=p3$.
\end{remark}

In Appendix \ref{pseudocode appendix}, we provide the full pseudocode of the WEST program.
We summarize the important points to provide analysis of time and space complexity. The regular expression $R = \{R_1 | R_2 | ... | R_k\}$, where for all $1 \leq i \leq k$, $R_i \in \{$`0', `1', `S', `,'$\}^*$, is represented as a vector $R = {R_1, R_2, ..., R_k}$.

In order to compute MLTL operations, we define functions for handling union and intersection of such vectors. 
The \textit{join} function performs simple concatenation of two vectors, while the \textit{set\_intersect} function calculates pairwise intersection of strings from two vectors.
The intersection of two strings is performed by \textit{bit\_wise\_and}, which computes intersection character by character.
For example, \textit{bit\_wise\_and}$(1s, 0s)$ is the empty string while \textit{bit\_wise\_and}$(s11, 0s1) = 011$.

The main focus is on the recursive function \textit{reg}, which is based on the structure of the grammar. It computes the propositional logic operators $>$ and  $=$ through formula rewriting in terms of And and Or. The four temporal operators in the grammar are computed according to the definitions of their regexes (See section \ref{regex}).

If the user would like to see the satisfying computations for all the subformulas in their original formula, \textit{reg} contains a Boolean to enable this functionality.

It is often the case that a vector of regular expression strings contains many redundancies, and can be simplified to a smaller vector of regular expressions that improves space complexity and drastically enhances user readability. 
As such, the \textit{simplify} function performs this by iterating through pairs of strings from the vector to search for valid simplifications. 
If a pair of strings$(w, v)$ can be written in the form $w = w_1 c_w w_2$ and $v = v_1 c_v v_2$ such that $w_1 = v_1$ and $w_2 = v_2$, then they are combined into one string $w_1 (c_w | c_v) w_2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Complexity} \label{complex}

\subsection{Theoretical Complexity}
 In order to reason about the complexity of our algorithms, we make a few reasonable assumptions about our inputs.
Suppose that the lower and upper intervals of temporal operators are bounded by some constant $d \in \mathbb{N}$, and that the difference between any bound is less than some constant $\delta \in \mathbb{N}$. 
We provide a summary of the complexity of each of the helper functions described for \textit{reg}. 
For any function $f(\phi)$ taking a string argument $\phi$, we use $|\phi|$ to denote the number of characters in $\phi$ and $S(f(\phi))$ to denote the space complexity of $f$ in terms of the number of characters in the output.\\

\textbf{If $\phi$ is a propositional constant or variable:}
%\begin{adjustwidth}{10mm}{3em}
    It's easy to see that $S(\text{\emph{reg}}(!)) = 0$ since only the empty vector is returned. \\
    By definition $\text{\emph{reg}}(T) = S^n$, so we have that $S(\text{\emph{reg}}(T)) = O(n)$. Similar to $\phi = T$, $\text{\emph{reg}}(p_k)$ and $\text{\emph{reg}}(\scriptstyle{\sim} \textstyle p_k)$ both return strings of the same length and thus we also have $S(\text{\emph{reg}}(p_k)) = S(\text{\emph{reg}}(\scriptstyle{\sim} \textstyle p_k)) = S(\text{\emph{reg}}(T)) = n$.\\
%\end{adjustwidth}

\textbf{If} $\phi$ \textbf{is} ``$\phi_{1} \lor \phi_{2}$'' \textbf{:}
%\begin{adjustwidth}{10mm}{3em}
    For this case in our \emph{reg} function, we return $\textit{join}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}))$ \\
    Thus, our space complexity is:
    \[
    S(\text{\emph{reg}}(\phi_{1} \lor \phi_{2})) = S(\text{\emph{reg}}(\phi_{1})) + S(\text{\emph{reg}}(\phi_{2})).
    \]
%\end{adjustwidth}

\textbf{If} $\phi$ \textbf{is} ``$\phi_{1} \; \& \; \phi_{2}$'' \textbf{:}
%\begin{adjustwidth}{10mm}{3em}
    For this case in our \emph{reg} function, we return $ \textit{set\_intersect}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}), n)$ \\
    In the worst case, \textit{set\_intersect}($A$, $B$, $n$) will return a vector of size $S(A) \cdot S(B)$.\\
    Thus, our space complexity is:
    \[ 
    S(\text{\emph{reg}}(\phi_{1} \; \& \; \phi_{2})) = S(\text{\emph{reg}}(\phi_{1})) \cdot S(\text{\emph{reg}}(\phi_{2})).
    \]
%\end{adjustwidth}

\textbf{If} $\phi$ \textbf{is} ``$\phi_{1} >  \phi_{2}$'' \textbf{:}
%\begin{adjustwidth}{10mm}{3em}
    For this case in our \emph{reg} function, we first rewrite $\phi_{1} > \phi_{2}$ to the equivalent formula:
    \textit{equiv\_formula} = \text{``(''} + \textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{1}) + \text{``} \lor \text{''} + \phi_{2} + \text{``)''}.
    Then we return $\text{\emph{reg}}(\textit{equiv\_formula})$.
    So for space complexity, we have:
    S(\text{\emph{reg}}(\phi_{1} > \phi_{2})) = S(\text{\emph{reg}}(\textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{1})) + S(\text{\emph{reg}}(\phi_{2})).
%\end{adjustwidth}
\textbf{If} $\phi$ \textbf{is} ``$\phi_{1} = \phi_{2}$'' \textbf{:}
%\begin{adjustwidth}{10mm}{3em}
    For this case in our \emph{reg} function, we first rewrite $\phi_{1} = \phi_{2}$ to the equivalent formula:
    \textit{equiv\_formula} = 
    \text{``((''} + \phi_{1} + \text{``} \& \text{''} + \phi_{2} + \text{``)}
    \lor 
    \text{(''} + \textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{1}) + \text{``} \& \text{''} +  \textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{2}) + \text{``))''}.
    Then we return $\text{\emph{reg}}(\textit{equiv\_formula})$.
    So for space complexity, we have:
    S(\text{\emph{reg}}(\phi_{1} = \phi_{2})) =
    S(\text{\emph{reg}}(\phi_{1})) \cdot S(\text{\emph{reg}}(\phi_{2})) \; + \;  
    S(\text{\emph{reg}}(\textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{1})) \cdot S(\text{\emph{reg}}(\textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{2})).
%\end{adjustwidth}
\textbf{If} $\phi$ \textbf{is} ``$\lor$ [ $\phi_{1}$, .., $\phi_{k}$]'' \textbf{:} 
%\begin{adjustwidth}{10mm}{3em}
For this case in our \emph{reg} function, we first rewrite $\lor [\phi_{1}, .., \phi_{k}]$ to the equivalent formula:
    \textit{equiv\_formula} = \text{``}( ... (\phi_{1} \lor \phi_{2}) \lor ... \phi_{k})\text{''}.
Using the space complexity of $\lor$, we have that:
    S(\text{\emph{reg}}(\lor [\phi_{1}, .., \phi_{k}])) = 
    \sum_{i=1}^{k} S(\text{\emph{reg}}(\phi_{i})).
%\end{adjustwidth}
\textbf{If} $\phi$ \textbf{is} ``$\&$ [ $\phi_{1}$, .., $\phi_{k}$]'' \textbf{:} 
%\begin{adjustwidth}{10mm}{3em}
For this case in our \emph{reg} function, we first rewrite $\& [\phi_{1}, .., \phi_{k}]$ to the equivalent formula:
    \textit{equiv\_formula} = \text{``}( ... (\phi_{1} \& \phi_{2}) \& ... \phi_{k})\text{''}.
Using the space complexity of $\&$, we have that:
    S(\text{\emph{reg}}(\& [\phi_{1}, .., \phi_{k}])) = 
    \prod_{i=1}^{k} S(\text{\emph{reg}}(\phi_{i})).
%\end{adjustwidth}
\textbf{If} $\phi$ \textbf{is} ``$=$ [$\phi_{1}$, .., $\phi_{k}$]'' \textbf{:}
%\begin{adjustwidth}{10mm}{3em}
For this case in our \emph{reg} function, we first rewrite $= [\phi_{1}, .., \phi_{k}]$ to the equivalent formula:
    \textit{equiv\_formula} = \text{``}( ... (\phi_{1} = \phi_{2}) = ... \phi_{k})\text{''}.
Based on the space complexities of $\&$ and $\lor$, this gives:
    S(\text{\emph{reg}}(= [\phi_{1}, .., \phi_{k}])) = 
    \prod_{i=1}^{k} S(\text{\emph{reg}}(\phi_{i}))
    + \prod_{i=1}^{k} S(\text{\emph{reg}}(\textit{wff\_to\_nnf}(\scriptstyle{\sim}\textstyle \phi_{i}))).
%\end{adjustwidth}
 \textbf{If} $\phi \text{ is } ``\mathcal{G}_{[a,b]} \phi_{1}$"\textbf{:}
% \begin{adjustwidth}{10mm}{3em}
 Recall that $\text{\emph{reg}}(\mathcal{G}_{[a,b]} \phi) = \bigwedge_{i=a}^{b} (S^n,)^i \text{\emph{reg}}(\phi) (,S^n)^{b-i}$.\\
 As shown in the \textit{Assoc\_Prop\_Conn} case, if $\phi$ is ``$\&[\phi_1, ... ,\phi_k$]", then \\
 $\text{S(\emph{reg}}(\& [\phi_{1}, .., \phi_{k}])) = 
    \prod_{i=1}^{k} \text{S(\emph{reg}}(\phi_{i}))$. We have that:
     \text{S(\emph{reg}}(\phi)) &= \prod_{i=a}^{b} (n+1)i \cdot \text{S(\emph{reg}}(\phi_{1}))\\
     &< \Bigl([(n+1)b+1] \cdot \text{S(\emph{reg}}(\phi_1))\Bigl)^\delta \\
     &= \mathcal{C}_{G} \cdot \text{S(\emph{reg}}(\phi_1)^\delta 
 where $\mathcal{C}_{G} = [(n+1)b+1]^{\delta}$.
 In the calculation, $(n+1)i$ counts the concatenation of the padded component of the computation in the worse case when we have a vector of only length 1 strings. \\
% \end{adjustwidth}
 \textbf{If} $\phi \text{ is } ``\mathcal{F}_{[a,b]} \phi_{1}$"\textbf{:}
% \begin{adjustwidth}{10mm}{3em}
 As shown in the \textit{Assoc\_Prop\_Conn} case, if $\phi$ is ``$\lor[\phi_1, ... ,\phi_k$]", then \\
 $S(\text{\emph{reg}}(\lor [\phi_{1}, .., \phi_{k}])) = 
    \sum_{i=1}^{k} S(\text{\emph{reg}}(\phi_{i}))$. Thus with similar calculation to $\mathcal{G}$,
     \text{S(\emph{reg}}(\phi)) &= \sum_{i=a}^{b} (n+1)i \cdot \text{S(\emph{reg}}(\phi_{1})) \\
     &< \delta\Bigl(((n+1)b+1) \cdot \text{S(\emph{reg}}(\phi_1))\Bigl) \\
     &= \mathcal{C}_{F} \cdot \text{S(\emph{reg}}(\phi_1)
 where $\mathcal{C}_{F} = \delta[(n+1)b+1]$.
% \end{adjustwidth}

 \textbf{If $\phi$ = $``\phi_1 \mathcal{U}_{[a, b]} \phi_2"$:}
% \begin{adjustwidth}{10mm}{3em}
    By our previous definitions, 
    $\text{\emph{reg}}(\phi_1 \ \mathcal{U}_{[a,b]} \phi_2) =  \bigvee_{i=a}^{b} \text{\emph{reg}}\left(\mathcal{G}_{[a,i-1]}\phi_1 \land \mathcal{G}_{[i, i]} \phi_2\right)$. \\
    We can bound $S(\text{\emph{reg}}(\mathcal{G}_{[i, i]} \phi_2))$ by $(n+1) \cdot i \cdot S(\text{\emph{reg}}(\phi_2))$ because the operation is equivalent to simply prepending $(S^n ,)^i$.
    Thus using our previous results for the $\mathcal{G}$, $\&$ and $\lor$ operators, we have that:
    \begin{align*}
        S(\text{\emph{reg}}(\phi)) &= 
        \sum_{i = a}^b [((n+1)(i-1) + 1) \cdot S(\text{\emph{reg}}(\phi_1))]^\delta 
        [(n+1) \cdot i \cdot S(\text{\emph{reg}}(\phi_2))] \\
        &< \delta [((n+1)(b-1) + 1) \cdot S(\text{\emph{reg}}(\phi_1))]^\delta 
        [(n+1) \cdot b \cdot S(\text{\emph{reg}}(\phi_2))] \\
        &< \delta ((n+1)(b-1) + 1)^\delta \cdot (n+1) \cdot d \cdot S(\text{\emph{reg}}(\phi_1))^\delta \cdot S(\text{\emph{reg}}(\phi_2)) \\
        &= \mathcal{C}_U \cdot S(\text{\emph{reg}}(\phi_1))^\delta \cdot S(\text{\emph{reg}}(\phi_2))
    \end{align*}
    where $\mathcal{C}_U = \delta ((n+1)(b-1) + 1)^\delta \cdot (n+1) \cdot d$.
%  \end{adjustwidth}

  \textbf{If $\phi$ = $``\phi_1 \mathcal{R}_{[a, b]} \phi_2"$:}
%  \begin{adjustwidth}{10mm}{3em}
        By our previous definitions, $\text{\emph{reg}}(\phi_1 \mathcal{R}_{[a,b]} \phi_2) =  \text{\emph{reg}}\left(\mathcal{G}_{[a,b]}\phi_2\right) \lor \bigvee_{i=a}^{b-1} \text{\emph{reg}}\left(\mathcal{G}_{[a,i]}\phi_2 \land \mathcal{G}_{[i, i]} \phi_1\right)$. Thus, a similar argument to the $\mathcal{U}$ case gives us that: 
        \begin{align*}
            S(\text{\emph{reg}}(\phi)) &< 
            \mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta + \sum_{i = a}^{b-1} [\mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta \cdot (n+1)\cdot i \cdot S(\text{\emph{reg}}(\phi_2))] \\
            &< \mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta + \delta[\mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta \cdot (n+1)(b-1) \cdot S(\text{\emph{reg}}(\phi_2))] \\
            &= \mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta [1 + \delta(n+1)(b-1)S(\text{\emph{reg}}(\phi_2))] \\
            &< \mathcal{C}_G \cdot S(\text{\emph{reg}}(\phi_1))^
            \delta (\delta(n+1)d \cdot S(\text{\emph{reg}}(\phi_2))) \\
            &= \mathcal{C}_R \cdot S(\text{\emph{reg}}(\phi_1))^\delta \cdot S(\text{\emph{reg}}(\phi_2))
        \end{align*}
        where $\mathcal{C}_R = \mathcal{C}_G \cdot \delta (n+1)d$. \\
%  \end{adjustwidth}
\begin{theorem}[\emph{Space Complexity}]
Given a well formed MLTL formula $\phi$, $\text{\emph{reg}}(\phi)$ has worst-case space complexity that is $O(\mathcal{C}_U^{\delta ^ \ell} \cdot (\ell + 1)^{\delta^{\ell + 1}})$, where $\ell$ is the number of logical connectives ($\land, \lor, \mathcal{F}, \mathcal{G}, \mathcal{R}, \mathcal{U}$) in $\phi$. 
\end{theorem}
\begin{proof}
    To analyze worst-case complexity, it's clear from the analysis above that $\mathcal{U}$ and $\mathcal{R}$ give the largest complexity. From the constants in their respective bounds, we can show that $\mathcal{C}_U = ((n+1)d - n)^\delta \cdot \mathcal{C}_R > \mathcal{C}_R$, and so we'll analyze only repeated nesting of the operator $\mathcal{U}$.

>>>>

%KYR: Take this out and replace it with a statement that we use the standard RE operators:...
%% \begin{definition}[\emph{Regular Expression \cite{sipser}}]
%% A \emph{regular expression} is a string that describes the form of a language. 
%% Let $\Sigma$ denote a finite phibet. The empty set $\emptyset$, the empty string $\epsilon = ``"$, and any character in $\Sigma$ are constants defined to be a regular expression. 
%% Beyond that, we define regular expressions recursively. 
%% Suppose $R$ and $T$ are regular expressions. 
%% Then the following operations generate regular expressions:
%% \begin{itemize}
%%     \item \emph{(Concatenation)} Denoted $RT$, this operation describes the set of strings obtained by concatenating any string generated by $R$ and any string generated by $T$ in that order.
%%     \item \emph{(Alternation)} Denoted $R | T$, this operation describes the set union of the strings generated by $R$ and $T$.
%%     \item \emph{(Kleene Star)} Denoted $R^*$, this operation describes the set of all strings made by concatenating any finite number (including zero) of strings described by $R$ together.
%%     \item $R^i$ denotes regular expression consisting of $R$ repeated $i$ times for $i \geq 0$. $R^0 = \epsilon$.
%% \end{itemize}
%% \end{definition}
%
%KYR: Examples are super nice, but if we have them we should have them after the MLTL regex example
%\noindent We provide an example of each operation.
%% \begin{example}
%% Let $R$ describe \{`$a$', `$b$'\} and $T$ describe \{ `$c$', `$d$' \}.
%% \begin{itemize}
%%     \item $RT$ describes the set \{ `$ac$', `$ad$', `$bc$', `$bd$' \}.
%%     \item $R | T$ describes the set \{`$a$', `$b$', `$c$', `$d$' \}.
%%     \item $R^*$ describes the set \{ $\epsilon$, `$a$', `$b$', `$aa$', `$bb$', `$ab$', `$ba$', `$aaa$', `$aba$',... \}.
%%     \item $R^2$ describes the set \{ `$aa$', `$bb$', `$ab$', `$ba$' \}.
%% \end{itemize}
%% \end{example}

%% %\noindent We provide examples of different regex and possible combinations of the above operations.
%% \begin{example}
%% Let `$a$', `$b$', and `$c$' denote characters in a finite phibet.
%% \begin{itemize}
%%     \item $ab^*$ describes the set \{ `$a$', `$ab$', `$abb$', `$abbb$', ... \}.
%%     \item $a^*|b$ describes the set \{ $\epsilon$, `$b$', `$a$', `$aa$', `$aaa$', ...\}.
%%     \item $a(b^*)c$ describes the set \{ `$ac$', `$abc$', `$abbc$', `$abbbc$', ... \}.
%%     \item $(a | b)^3$ describes the set \{ `$aaa$', `$aba$', `$aab$', `$abb$', `$bbb$', ... \}.
%% \end{itemize}
%% \end{example}

@@

 \end{itemize} % \vspace{-0.1in} % \subsection{Soundness and Completeness} %  \vspace{-0.1in}
 We write $\mathscr{L}(\text{reg}(\phi))$ to denote the language of reg($\phi$). That is, the set of computations represented by the regular expression $\text{reg}(\phi)$. % The proof of Theorem \ref{SoundComplete} appears in Appendix \ref{Appendix-SoundComplete}.

@@

%%  \noindent This completes the inductive step, and thus the proof. Since this proof addresses all possible MLTL formulas in negation normal form, it shows completeness along with soundness.
%%  \end{proof}
% \newpage
%\vspace{-0,1in}
%\subsection{Nested Until and Release Rewriting Theorem}
%\vspace{-0,1in}
 We prove the correctness of our formula rewriting algorithm using our regular expressions from Definition \ref{TRE}. As detailed in Section \ref{complex}, nested binary temporal operators create the worst-case time and space complexity for the WEST program. Although the WEST program does not implement this theorem by default, we would expect selected applications of rewriting to improve runtime. We use this theorem to simplify MLTL formulas, making them easier to interpret.
% Theorem \ref{NestedUR}'s proof appears in Appendix \ref{Appendix-NestedUR}.

@@

\section{WEST Algorithm and Analysis} \label{WESTalg}
We now present the  full WEST algorithm, which takes MLTL formula $\phi$ (in NNF) and number of propositional variables $n$, and recursively computes all satisfying temporal regular expressions to $\phi$. 
We use C++ vectors of strings to represent alternation of regular expressions; for instance $\{t_1, t_2, ..., t_n\} = \bigcup_{i=1}^n t_i$. 
Additionally, we provide details for performing set intersection of temporal regular expressions, and the algorithm for the temporal operator $\mathcal{F}$. 
Note how the algorithm parallels the regular expression defined for $\mathcal{F}$, and the algorithms for the other three temporal operators follow an identical structure.

One major detail that drastically improves running time is applying a simplifying function to the output of set intersection. 
This is because the set intersection operation is repeatedly used to compute the temporal operators $\mathcal{G}, \mathcal{U}, \mathcal{R}$. 
Although set intersection does blow up exponentially in space in the worse case, it rarely does in practice; thus why we frequently simplify outputs. 
We simplify alternation by using the fact that for regular expressions $w_1$ and $w_2$, $w_1 1 w_2 \lor w_1 0 w_2 = w_1 s w_2$, and check for the reduction between at every index between every pair of strings in a given vector.

% \begin{algorithm}[H]
% \caption{WEST Algorithm}
% \begin{algorithmic}[1]
%     \Procedure{reg}{string $\phi$, int $n$}
%         \If{$\phi$ is $\top$ or $\bot$} return reg\_prop\_const($\phi$, n)
%         \EndIf

%         \If{$\phi$ is $p_k$ or $\neg p_k$} return reg\_prop\_var($\phi$, n)
%         \EndIf

%         \If{$\phi  = \phi_1 \land \phi_2$} return set\_intersect(reg($\phi_1$), reg($\phi_2$), n)
%         \EndIf

%         \If{$\phi  = \phi_1 \lor \phi_2$} return join(reg($\phi_1$), reg($\phi_2$), n)
%         \EndIf

%         \If{$\phi = \mathcal{F}_{[a, b]}\phi_1$} return reg\_F(reg($\phi_1$), a, b, n)
%         \EndIf

%         \If{$\phi = \mathcal{G}_{[a, b]}\phi_1$} return reg\_G(reg($\phi_1$), a, b, n)
%         \EndIf

%         \If{$\phi = \phi_1 \mathcal{U}_{[a, b]} \phi_2$} return reg\_U(reg($\phi_1$), reg($\phi_2$), a, b, n)
%         \EndIf

%         \If{$\phi = \phi_1 \mathcal{R}_{[a, b]} \phi_2$} return reg\_R(reg($\phi_1$), reg($\phi_2$), a, b, n)
%         \EndIf
%     \EndProcedure
% \end{algorithmic}
% \end{algorithm}

% ZILI: Presenting algorithm vertically instead
\hspace{-5mm}
\begin{minipage}{0.52\textwidth}
\begin{algorithm}[H]
\caption{WEST Algorithm}
\begin{algorithmic}[1]
    \Procedure{reg}{string $\phi$, int $n$}
        \If{$\phi$ is $\top$ or $\bot$}
            \State return reg\_prop\_const($\phi$, n)
        \EndIf

        \If{$\phi$ is $p_k$ or $\neg p_k$}
            \State return reg\_prop\_var($\phi$, n)
        \EndIf

        \If{$\phi  = \phi_1 \land \phi_2$}
            \State \parbox[t]{.82\linewidth}{return set\_intersect(reg($\phi_1$), reg($\phi_2$), \hspace*{5mm}n)}
        \EndIf

        \If{$\phi  = \phi_1 \lor \phi_2$}
            \State return join(reg($\phi_1$), reg($\phi_2$), n)
        \EndIf

        \If{$\phi = \mathcal{F}_{[a, b]}\phi_1$}
            \State return reg\_F(reg($\phi_1$), a, b, n)
        \EndIf

        \If{$\phi = \mathcal{G}_{[a, b]}\phi_1$}
            \State return reg\_G(reg($\phi_1$), a, b, n)
        \EndIf

        \If{$\phi = \phi_1 \mathcal{U}_{[a, b]} \phi_2$}
            \State return reg\_U(reg($\phi_1$), reg($\phi_2$), a, b, n)
        \EndIf

        \If{$\phi = \phi_1 \mathcal{R}_{[a, b]} \phi_2$}
            \State return reg\_R(reg($\phi_1$), reg($\phi_2$), a, b, n)
        \EndIf
    \EndProcedure
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{algorithm}[H]
\caption{Takes set intersection of the languages of two regexes}
%Input: Vectors of strings that represent regexes, simplify boolean\\
%Output: Vector of strings that represents set intersection of the languages of the inputted regexes
\begin{algorithmic}
\Procedure{set\_intersect}{vector $R$, vector $T$, bool simp}
    \State Pad($R$, $T$), return $\leftarrow \{\}$
    \For{$(r, t) \in R\times T$}
            \State add bit\_wise\_and($r, t$) to ret 
    \EndFor
    \State return simplify(ret)
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Computes the regex for a MLTL formula $\phi$U[a:b]$\psi$} 
%Inputs: Vectors of strings representing the regexes for $\phi$ and $\psi$, interval bounds, number of propositional variables, simplify boolean\\
%Output: Vector of strings that represents the appropriate satisfying computations 
\begin{algorithmic}
\Procedure{reg\_U}{vector reg\_phi, vector reg\_psi int $a$, int $b$, int $n$, bool simp} 
    \State comp $\leftarrow$ ((`$S$')$^{n}$ + `,')$^{a} + $ reg\_psi
    \If{$a > b$} return $\{\}$
    \EndIf
    \For{($a \leq i \leq b-1$) }
        \State temp\_G1 $\leftarrow$ reg\_G(reg\_phi, $a$, $i$, $n$)
        \State \parbox[t]{.8\linewidth}{temp\_G2 $\leftarrow$ reg\_G(reg\_psi, $i+1$, \hspace*{5mm}$i+1$, $n$)}
        \State \parbox[t]{.8\linewidth}{temp\_comp $\leftarrow$ set\_intersect( \hspace*{5mm}temp\_G1, temp\_G2, $n$, simp)}
        \State \parbox[t]{.8\linewidth}{comp $\leftarrow$ join( comp, temp\_comp, \hspace*{5mm}simp)}
    \EndFor
    \State return comp
\EndProcedure
\end{algorithmic}
\end{algorithm}
\end{minipage}
